### 2017-04-17 paper/oopsla17-submission/conclusion.tex
re-add a summary sentence to conclusion

We have presented \toolname, which
combines modern statistical methods
with domain-specific feature engineering
to open the door to a new data-driven
path towards precise error localization,
significantly outperforming the
state of the art on a new benchmark
suite comprising 4,500 student programs.
%
%%% RJNUKEDME   \toolname
%%% RJNUKEDME   translates pairs of
%%% RJNUKEDME   blame-labeled ill-typed
%%% RJNUKEDME   programs into bags-of-abstracted-terms,
%%% RJNUKEDME   and then trains a classifier
%%% RJNUKEDME   to predict whether a feature vector
%%% RJNUKEDME   should get a blame label.
%%% RJNUKEDME   % that
%%% RJNUKEDME   % predicts the value of the
%%% RJNUKEDME   % whether a feature vector gets
%%% RJNUKEDME   % a blame label.
%%% RJNUKEDME   %
%%% RJNUKEDME   Given a new ill-typed program,
%%% RJNUKEDME   we extract its bag-of-abstracted-terms
%%% RJNUKEDME   to compute the blame label for each
%%% RJNUKEDME   term and rank the predictions based
%%% RJNUKEDME   on the model's confidence, to select
%%% RJNUKEDME   the top 1--3 blame assignments to
%%% RJNUKEDME   present to the user.
%%% RJNUKEDME
%%% RJNUKEDME   We have evaluated our approach with
%%% RJNUKEDME   five different classifiers on a suite
%%% RJNUKEDME   of over 4,500 student programs.
%%% RJNUKEDME   %
%%% RJNUKEDME   We found that while machine learning
%%% RJNUKEDME   over syntactic features of each term in isolation
%%% RJNUKEDME   performs worse than existing
%%% RJNUKEDME   purely constraint based approaches, % (\eg \ocaml, \sherrloc),
%%% RJNUKEDME   augmenting the data with a single feature corresponding to
%%% RJNUKEDME   the type error slice brings our
%%% RJNUKEDME   classifiers up to par with the state of the art,
%%% RJNUKEDME   and further augmenting the data with
%%% RJNUKEDME   features of an expression's parent and children
%%% RJNUKEDME   allows our classifiers to outperform
%%% RJNUKEDME   the state of the art by 18 percentage points.

>>>>

We have presented \toolname, which
combines modern statistical methods
with domain-specific feature engineering
to open the door to a new data-driven
path towards precise error localization,
significantly outperforming the
state of the art on a new benchmark
suite comprising 4,500 student programs.
%
%%% RJNUKEDME   \toolname
%%% RJNUKEDME   translates pairs of
%%% RJNUKEDME   blame-labeled ill-typed
%%% RJNUKEDME   programs into bags-of-abstracted-terms,
%%% RJNUKEDME   and then trains a classifier
%%% RJNUKEDME   to predict whether a feature vector
%%% RJNUKEDME   should get a blame label.
%%% RJNUKEDME   % that
%%% RJNUKEDME   % predicts the value of the
%%% RJNUKEDME   % whether a feature vector gets
%%% RJNUKEDME   % a blame label.
%%% RJNUKEDME   %
%%% RJNUKEDME   Given a new ill-typed program,
%%% RJNUKEDME   we extract its bag-of-abstracted-terms
%%% RJNUKEDME   to compute the blame label for each
%%% RJNUKEDME   term and rank the predictions based
%%% RJNUKEDME   on the model's confidence, to select
%%% RJNUKEDME   the top 1--3 blame assignments to
%%% RJNUKEDME   present to the user.
%%% RJNUKEDME
%%% RJNUKEDME   We have evaluated our approach with
%%% RJNUKEDME   five different classifiers on a suite
%%% RJNUKEDME   of over 4,500 student programs.
%
We found that while machine learning
over syntactic features of each term in isolation
performs worse than existing
purely constraint based approaches, % (\eg \ocaml, \sherrloc),
augmenting the data with a single feature corresponding to
the type error slice brings our
classifiers up to par with the state of the art,
and further augmenting the data with
features of an expression's parent and children
allows our classifiers to outperform
the state of the art by \ToolnameWinSherrloc
percentage points.

### 2017-04-17 paper/oopsla17-submission/evaluation-utility-graph.tex
forgot to account for list-literal feature
re-add a summary sentence to conclusion

\begin{subfigure}[t]{\linewidth}
\begin{tikzpicture}
\begin{axis}[
  % name=feature,
  % scale only axis,
  ybar stacked,
  width=0.9\linewidth,
  height=5cm,
  %title={Impact of Contextual Feature Classes},
  ylabel={Accuracy},
  bar width=0.5cm,
  ymin=0.5,
  ymax=1,
  ytick={0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  symbolic x coords={op, op+size, op+context, op+type, op+context+size, op+type+size, op+context+type, op+context+type+size},
  % enlarge x limits=0.5,
  xtick=data,
  xtick style={draw=none},
  xticklabel style={align=center},
  xticklabels={
    \textsc{Local Syntax}\\(44),
    \textsc{+Size}\\(45), \textsc{+Context}\\(220), \textsc{+Type}\\(102),
    +C+S\\(221), +T+S\\(103), +C+T\\(275),
    +C+T+S\\(276)
    % + Size, + Context, + Type,
    % + Context + Size, + Type + Size, + Context + Type,
    % + Context + Type + Size
  },
  %x tick label style={rotate=45, anchor=north east},
  % reverse legend,
  % transpose legend,
  % legend style={
  %   anchor = south east,
  %   at = {(1,1)},
  %   %legend pos = outer north east,
  %   legend columns=4},
]

@@

\end{tikzpicture}
\caption{Impact of contextual features on blame accuracy. Starting from
  a baseline of local syntactic features, we add each combination of
  expression size, contextual syntactic, and typing features. The total
  number of features is given in parentheses.}
\label{fig:context-utility}
\end{subfigure}

>>>>

\begin{subfigure}[t]{\linewidth}
\begin{tikzpicture}
\begin{axis}[
  % name=feature,
  % scale only axis,
  ybar stacked,
  width=0.9\linewidth,
  height=5cm,
  %title={Impact of Contextual Feature Classes},
  ylabel={Accuracy},
  bar width=0.5cm,
  ymin=0.5,
  ymax=1,
  ytick={0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
  yticklabel={\pgfmathparse{\tick*100}\pgfmathprintnumber{\pgfmathresult}\,\%},
  ytick style={draw=none},
  ymajorgrids = true,
  symbolic x coords={op, op+size, op+context, op+type, op+context+size, op+type+size, op+context+type, op+context+type+size},
  % enlarge x limits=0.5,
  xtick=data,
  xtick style={draw=none},
  xticklabel style={align=center},
  xticklabels={
    \textsc{Local Syntax}\\(45),
    \textsc{+Size}\\(46), \textsc{+Context}\\(225), \textsc{+Type}\\(100),
    +C+S\\(226), +T+S\\(101), +C+T\\(281),
    +C+T+S\\(282)
    % + Size, + Context, + Type,
    % + Context + Size, + Type + Size, + Context + Type,
    % + Context + Type + Size
  },
  %x tick label style={rotate=45, anchor=north east},
  % reverse legend,
  % transpose legend,
  % legend style={
  %   anchor = south east,
  %   at = {(1,1)},
  %   %legend pos = outer north east,
  %   legend columns=4},
]

@@

\end{tikzpicture}
\caption{
  %
  Impact of contextual features on blame accuracy.
  %
  % Starting from a baseline of local syntactic features,
  % we add each combination of
  % expression size, contextual syntactic, and typing features.
  %
  The total number of features for each set is given in parentheses.
}
\label{fig:context-utility}
\end{subfigure}

### 2017-04-17 paper/oopsla17-submission/related.tex
re-add a summary sentence to conclusion

\mypara{Explaining Type Errors}
In this paper we have focused solely on the task of \emph{localizing} a
type error, but a good error report should also \emph{explain} the
error.
%
\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
attempt to explain type errors by collecting the chain of inferences
made by the type checker %--- essentially the typing derivation ---
and presenting them to the user.
%
% Such explanations can be lengthy, as an attempt to compress the
% explanation \citet{Yang2000-kz} presents a visualization of the
% inference process.
%
\citet{Gast2004-zd} produces a slice enhanced by arrows
showing the dataflow from sources with different types to a
shared sink, borrowing the insight of dataflows-as-explanations from
\textsc{MrSpidey}~\citep{Flanagan1996-bu}.
%
\citet{Hage2006-hc} catalog a set of heuristics for
improving the quality of error messages by examining errors made by
novices.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} extend the ability to customize error messages to
library authors, enabling \emph{domain-specific} errors.
%
%\mypara{Interactive Explanations}
Such \emph{static} explanations of type errors run the risk of
overwhelming the user with too much information, it may be preferable to
treat type error diagnosis as an \emph{interactive} debugging session.
%
\citet{Bernstein1995-yj} extend the type inference procedure to handle
\emph{open} expressions (\ie with unbound variables), allowing users to
interactively query the type checker for the types of sub-expressions.
%
\citet{Chitil2001-td} proposes \emph{algorithmic debugging} of type
errors, presenting the user with a sequence of yes-or-no questions about
the inferred types of sub-expressions that guide the user to a specific
explanation for the error.
%
\citet{Seidel2016-ul} explain type errors by searching for inputs that
expose the \emph{run-time} error that the type system prevented, and
present users with an interactive visualization of the erroneous
computation.

@@

\mypara{Fault Localization}
%
Given a defect, \emph{fault localization} is the task of identifying
``suspicious'' program elements (\eg lines, statements) that are likely
implicated in the defect %(\ie that should be changed to fix the defect)
%
--- thus, type error localization can be viewed as an instance of fault
localization.
%
The best-known fault localization technique is likely Tarantula, which
uses a simple mathematical formula based on measured information from
dynamic normal and buggy runs~\citep{Jones2002-us}.
%
Other similar approaches, including those of \citet{Chen2002-qz} and
\citet{Abreu2006-fn,Abreu2007-mu} consider alternate features of
information or refined formulae and generally obtain more precise
results; see \citet{Wong2009-pd} for a survey.
%
While some researchers have approached such fault localization with an
eye toward optimality (\eg \citet{Yoo2013-rw} determine optimal
coefficients), in general such fault localization approaches are limited
by their reliance on either running tests or including relevant
features.
%
For example, Tarantula-based techniques require a normal and a buggy run
of the program.
%
By contrast, we consider incomplete programs with type errors that may
not be executed in any standard sense.
%
Similarly, the features available influence the classes of defects that
can be localized.
%
For example, a fault localization scheme based purely on control flow features
will have difficulty with cross-site scripting or SQL code injection
attacks, which follow the same control flow path on normal and buggy
runs (differing only in the user-supplied data).
%
Our feature set is comprised entirely of syntactic and typing features,
a natural choice for modeling type errors, but it would likely not
generalize to other types of defects.

>>>>

\mypara{Explaining Type Errors}
In this paper we have focused solely on the task of \emph{localizing} a
type error, but a good error report should also \emph{explain} the
error.
%
\citet{Wand1986-nw}, \citet{Beaven1993-hb}, and \citet{Duggan1996-by}
attempt to explain type errors by collecting the chain of inferences
made by the type checker %--- essentially the typing derivation ---
and presenting them to the user.
%
% Such explanations can be lengthy, as an attempt to compress the
% explanation \citet{Yang2000-kz} presents a visualization of the
% inference process.
%
\citet{Gast2004-zd} produces a slice enhanced by arrows
showing the dataflow from sources with different types to a
shared sink, borrowing the insight of dataflows-as-explanations from
\textsc{MrSpidey}~\citep{Flanagan1996-bu}.
%
\citet{Hage2006-hc} catalog a set of heuristics for
improving the quality of error messages by examining errors made by
novices.
%
\citet{Heeren2003-db}, \citet{Christiansen2014-qc}, and
\citet{Serrano2016-oo} extend the ability to customize error messages to
library authors, enabling \emph{domain-specific} errors.
%
%\mypara{Interactive Explanations}
Such \emph{static} explanations of type errors run the risk of
overwhelming the user with too much information, it may be preferable to
treat type error diagnosis as an \emph{interactive} debugging session.
%
\citet{Bernstein1995-yj} extend the type inference procedure to handle
\emph{open} expressions (\ie with unbound variables), allowing users to
interactively query the type checker for the types of sub-expressions.
%
\citet{Chitil2001-td} proposes \emph{algorithmic debugging} of type
errors, presenting the user with a sequence of yes-or-no questions about
the inferred types of sub-expressions that guide the user to a specific
explanation.
%
\citet{Seidel2016-ul} explain type errors by searching for inputs that
expose the \emph{run-time} error that the type system prevented, and
present users with an interactive visualization of the erroneous
computation.

@@

\mypara{Fault Localization}
%
Given a defect, \emph{fault localization} is the task of identifying
``suspicious'' program elements (\eg lines, statements) that are likely
implicated in the defect %(\ie that should be changed to fix the defect)
%
--- thus, type error localization can be viewed as an instance of fault
localization.
%
The best-known fault localization technique is likely Tarantula, which
uses a simple mathematical formula based on measured information from
dynamic normal and buggy runs~\citep{Jones2002-us}.
%
Other similar approaches, including those of \citet{Chen2002-qz} and
\citet{Abreu2006-fn,Abreu2007-mu} consider alternate features of
information or refined formulae and generally obtain more precise
results; see \citet{Wong2009-pd} for a survey.
%
While some researchers have approached such fault localization with an
eye toward optimality (\eg \citet{Yoo2013-rw} determine optimal
coefficients), in general such fault localization approaches are limited
by their reliance on either running tests or including relevant
features.
%
For example, Tarantula-based techniques require a normal and a buggy run
of the program.
%
By contrast, we consider incomplete programs with type errors that may
not be executed in any standard sense.
%
Similarly, the features available influence the classes of defects that
can be localized.
%
For example, a fault localization scheme based purely on control flow features
will have difficulty with cross-site scripting or SQL code injection
attacks, which follow the same control flow path on normal and buggy
runs (differing only in the user-supplied data).
%
Our feature set is comprised entirely of syntactic and typing features,
a natural choice for type errors, but it would likely not
generalize to other defects.
